Computer-Assisted Analysis of the \\ Anderson-Hájek Ontological Controversy

### 2015-07-14 Papers/InProgress/HajekAndersonControversy/main.tex
Add formalization of Anderson's simplification.

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[H:A12] The negation of a property necessarily implied by a
  positive property is not positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property:
  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$
\item[A3']  The property of being God-like is positive:
  $$\hol{P(G_H)}$$
\item[A4]  Positive properties are necessarily positive:
  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$
\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$
\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
  \exq y [\varphi(y)]]}$$
\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$

@@

Additionally, Anderson \citep[footnote
14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only
the quantifiers in T3' and in A:D2 need to be interpreted as
actualistic quantifiers, while others may be taken as possibilistic
quantifiers. Our computer-assisted study of this mixed variant shows
that A4 is still redundant in logic \KFourB, but A5' becomes independent (hence not
redundant). Unfortunately, a countermodel for T3 can then be found.

The controversy over the superfluousness of A4 and A5 indicates a
trend to reduce the ontological argument to its bare essentials. In
this regard,  already
\citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates
that, by taking a notion of \emph{defective} as primitive and defining
the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become
derivable. These claims have been confirmed by the automated theorem
provers (in logic \KFourB). Within the same trend, the alternative
proposed by \citet{bjordal99} (cf. Fig. \ref{fig:Bjordal})
achieves a high level of minimality.

@@

\begin{itemize}
\item[B:D1] A property is positive iff it is necessarily possessed by every God-like being.
  $$\hol{P_B(\phi) \equiv \nec \allq x (G_B(x) \imp \phi(x))}$$
\item[B:L1] B:D1 is logically equivalent in \SFour with the union of D1' and axioms A2', A3' and A4'.
  $$\hol{\textrm{B:D1} \biimp \textrm{D1'} \wedge \textrm{A2'} \wedge \textrm{A3'} \wedge \textrm{A4'} }$$
%  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
\item[B:D2] a \emph{maximal composite} of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
  $$\hol{\MCP(\phi,x) \equiv (\phi(x) \wedge P_B(\phi)) \wedge \allq \psi ((\psi(x) \wedge P_B(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
\item[B:D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its maximal composites.
  $$\hol{\NE_B(x) \equiv \allq \phi (\MCP(\phi,x) \imp \nec \exq y \phi(y))}$$
\item[A:A1'] If a property is positive, its negation is not positive:
  $$\hol{\allq \varphi [P_B(\varphi) \imp \neg P_B(\neg \varphi)]}$$
\item[A5'] Necessary existence is a positive property.
 $$\hol{P_B(\NE_B)}$$
\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_B(x)}$$
\end{itemize}
\end{minipage}
} %\vskip-.5em
\caption{Bjordal's Alternative \label{fig:Bjordal}}
\end{figure}

>>>>

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[H:DP] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property:
  $$\hol{P(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$
\item[A3']  The property of being God-like is positive:
  $$\hol{P(G_H)}$$
\item[A4]  Positive properties are necessarily positive:
  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$
\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$
\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences:
  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
  \exq y [\varphi(y)]]}$$
\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$

@@

Additionally, Anderson \citep[footnote
14]{anderson90:_some_emend_of_goedel_ontol_proof} (cf. Fig. \ref{fig:Anderson_simp}) remarks that only
the quantifiers in T3' and in A:D2 need to be interpreted as
actualistic quantifiers, while others may be taken as possibilistic
quantifiers. Our computer-assisted study of this mixed variant shows
that A4 is still redundant in logic \KFourB, but A5' becomes independent (hence not
redundant). Unfortunately, a countermodel for T3 can then be found.

The controversy over the superfluousness of A4 and A5 indicates a
trend to reduce the ontological argument to its bare essentials. In
this regard,  already
\citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates
that, by taking a notion of \emph{defective} as primitive and defining
the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become
derivable.

\begin{figure}[t]
\noindent \framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
$D$ (Defective) is taken as primitive and $P_{AS}$ (Positive) is defined.

\begin{itemize}
\item[AS:D1] A property is positive iff its absence necessarily renders an individual defective and it is possible
  that an individual has the property without being defective.
   $$\hol{P_{AS}(\varphi) \equiv \nec (\allq x (\neg \varphi(x) \imp D(x)) \wedge (\neg \nec \allq x (\varphi(x) \imp D(x))))}$$
\item[A:A1'] If a property is positive, its negation is not positive:
  $$\hol{\allq \varphi [P_{AS}(\varphi) \imp \neg P_{AS}(\neg \varphi)]}$$
\item[A2'] A property necessarily implied by a
  positive property is positive:
  $$\hol{\allq \varphi \allq \psi [(P_{AS}(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp P_{AS}(\psi)]}$$
\item[A4']  Positive properties are necessarily positive:
  $$\hol{\allq \varphi [P_{AS}(\varphi) \imp \nec P_{AS}(\varphi)]}$$
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Anderson's Simplification \label{fig:Anderson_simp}}
\end{figure}

These claims have been confirmed by the automated theorem
provers (in logic \KFourB). Within the same trend, the alternative
proposed by \citet{bjordal99} (cf. Fig. \ref{fig:Bjordal})
achieves a high level of minimality.

@@

\begin{itemize}
\item[B:D1] A property is positive iff it is necessarily possessed by every God-like being.
  $$\hol{P_B(\phi) \equiv \nec \allq x (G_B(x) \imp \phi(x))}$$
\item[B:L1] B:D1 is logically equivalent in \SFour with the union of D1' and axioms A2', A3' and A4'.
  $$\hol{\textrm{B:D1} \biimp \textrm{D1'} \wedge \textrm{A2'} \wedge \textrm{A3'} \wedge \textrm{A4'} }$$
%  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
\item[B:D2] a \emph{maximal composite} of an individual's positive properties is a positive property possessed by the individual and necessarily implying every positive property possessed by the individual.
  $$\hol{\MCP(\phi,x) \equiv (\phi(x) \wedge P_B(\phi)) \wedge \allq \psi ((\psi(x) \wedge P_B(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
\item[B:D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its maximal composites.
  $$\hol{\NE_B(x) \equiv \allq \phi (\MCP(\phi,x) \imp \nec \exq y \phi(y))}$$
\item[A:A1'] If a property is positive, its negation is not positive:
  $$\hol{\allq \varphi [P_B(\varphi) \imp \neg P_B(\neg \varphi)]}$$
\item[A5'] Necessary existence is a positive property.
 $$\hol{P_B(\NE_B)}$$
\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_B(x)}$$
\end{itemize}
\end{minipage}
} \vskip-.5em
\caption{Bjordal's Alternative \label{fig:Bjordal}}
\end{figure}
